How to select predictive models for decision-making or causal inference?

### 2023-12-16 how_to_select_causal_models.tex
iter

    \caption[The best predictor may not estimate best causal
        effects]{\textbf{Illustration}: a) a random-forest estimator
        with high performance for standard prediction (high $R^2$) but that
        yields poor causal estimates (large error between true effect $\tau$ and
        estimated $\hat{\tau}$), b) a linear estimator with smaller
        prediction performance leading to better causal estimation. \\[1ex]
        Selecting the estimator with the smallest error to the individual
        treatment effect $\mathbb{E}[(\tau(x) - \hat{\tau}(x))^2]$
        --the $\tau\text{-risk}$, eq.\,\ref{eq:tau_risk} -- would lead to
        the best causal estimates; however computing this error is not
        feasible: it requires access to unknown quantities:
        $\tau(x)$. \\[1ex]
        While the random forest fits the data better than the linear model, it
        gives worse causal inference because its error is inhomogeneous between
        treated and untreated. The $R^2$ score does not capture this
        inhomogeneity.
        \label{fig:toy_example}
    }
\end{figure}

>>>>

    \caption[The best predictor may not estimate best causal
        effects]{\textbf{Illustration}: a) a random-forest predictor
        with high performance for standard prediction (high $R^2$) but that
        yields poor causal estimates (large error between true effect $\tau$ and
        estimated $\hat{\tau}$), b) a linear predictor with smaller
        prediction performance leading to better causal estimation. \\[1ex]
        Selecting the predictor with the smallest error to the individual
        treatment effect $\mathbb{E}[(\tau(x) - \hat{\tau}(x))^2]$
        --the $\tau\text{-risk}$, eq.\,\ref{eq:tau_risk} -- would lead to
        the best causal estimates; however computing this error is not
        feasible: it requires access to unknown quantities:
        $\tau(x)$. \\[1ex]
        While the random forest fits the data better than the linear model, it
        gives worse causal inference because its error is inhomogeneous between
        treated and untreated. The $R^2$ score does not capture this
        inhomogeneity.
        \label{fig:toy_example}
    }
\end{figure}
